---
title: "Env Econ Assignmnet 1"
author: Travis Rennacker
format:
  html: 
   toc: true
   number-sections: true
   code-fold: true
execute: 
  error: false
---

Quarto libraries

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(here)
library(pracma)
        
```

1a,1b

```{r}
z <- seq(1,10, by=0.5) 

z

liney <- function(a,x,b){
   y <- (a*x+b)
   return(y)
}

liney(3,5,10)

```

1c

```{r}
df <- data.frame(x = seq(-1,10))

a<-3
b<-10

df$a <- a
df$b <- b
df$y <- df$a*df$x+df$b

df


```

1d

```{r}
ggplot(data=df, aes(x=x, y=y))+geom_line()
```

1e

```{r}
rm(list = ls(all = TRUE))  

a <- -2
b <- 100
s <- 3
d <- 0

func1 <- function(a,b,X){
  Y=a*X+b
  return(Y)
}

func2 <- function(s,d,X){
  Y=s*X+d
  return(Y)
}

func3 <- function(x,s,a,b,d){
  Z = func1(a,b,x) - func2(s,d,x)
  return(Z)
}


rooty_1 <- fzero(func3, a = -2,
b = 100, s = 3, d = 0, c(-100, 100))


rooty_1$x

the_fore_told_Y_value_at_x_20 <- func1(-2,100,20)

the_fore_told_Y_value_at_x_20


```

# 1f : algebra confirms all answers in above part 1e

2

```{r}
 
pnut_demand <- function(q){
  p=500-0.1*q
  return(p)
}
  
pnut_supply <- function(q){
  p=5+0.05*q
  return(p)
}

eq_pnut <- function(q){
  
  z=pnut_demand(q)-pnut_supply(q)
  return(z)
}

qstar <- fzero(eq_pnut, c(-100,10000))$x

price_pnut <- pnut_demand(3300)

qstar
price_pnut
```

2a: equilibrium price will be at 170 credits at quantity of 3300 jars

```{r}
dq_val <- seq(-1,5000, by=1)
dp_val <- 500-0.1*dq_val

demand_pnut <- data.frame(x=dq_val, y=dp_val)

sq_val <- seq(-1,10000)
sp_val <- 5+0.05*sq_val

supply_pnut <- data.frame(x=sq_val, y=sp_val)

ggplot() + geom_line(data=demand_pnut, aes(x=x, y=y, color="Demand")) + geom_line(data=supply_pnut, aes(x=x,y=y, color="Supply")) +labs(x="quantity", y= "price")


```

2b Benefit

```{r}
supply_benefit <- (qstar*price_pnut)/2

supply_benefit

demand_benefit <- (3300*(500-170))/2

demand_benefit 
```

2c New New York Large Food Bank

```{r}
dnyq_val <- seq(-1,200, by=1)
dnyp_val <- 400-2*dnyq_val

ny_demand <- data.frame(x=dnyq_val, y=dnyp_val)
  
  ggplot() + geom_line(data=demand_pnut, aes(x=x, y=y, color="Demand")) + geom_line(data=supply_pnut, aes(x=x,y=y, color="Supply")) + geom_line(data=ny_demand, aes(x=x,y=y, color="NY Demand")) +labs(x="quantity", y= "price")
```

2c

```{r}
agq_val <- seq(-1,200)
agp_val <- 900-2.1*agq_val

ag_demand <- data.frame(x=agq_val, y=agp_val)

final_dq_val <- seq(200,5000, by=1)
final_dp_val <- 500-0.1*final_dq_val

final_ag_demand <- data.frame(x=final_dq_val, y=final_dp_val)

 ggplot() + geom_line(data=final_ag_demand, aes(x=x, y=y, color="Demand")) + geom_line(data=ag_demand, aes(x=x, y=y, color="Demand")) + geom_line(data=supply_pnut, aes(x=x,y=y, color="Supply")) + labs(x="quantity", y= "price")


```

2c The new NY Food Bank does not effect the equilibrium price. The equilibrium price remains at 170 credits.

2d If trading jars of peanut butter generates carbon emissions than these emissions should be incorporated into the cost of the trade. Would this mean that the carbon emissions would not become an "externality?"

3a used car market supply p=200+0.2\*q\^1.2

```{r}
rm(list = ls(all = TRUE))  

quantity <- seq(0,2000)
p <- 2000

CA_price <- data.frame(x=quantity, y=p)

quantity1 <- seq(0,2000)
p1 <- 200+0.2*(quantity1^1.2)

ucar_sup <- data.frame(x=quantity1, y= p1)

CA_purchase <- function(q){
  return(2000)
}

sup <- function(q) {
  p=200+0.2*(q^1.2)
  return(p)
}

equ <- function(q) {
  z=sup(q)-CA_purchase(q)
  return(z)
}

quantity_cars <- fzero(equ, c(0,3000))$x

quantity_cars


```

3a i.1973 cars will be bought by the state given this supply curve

ii. Elasticity of supply at equilibrium e=(dP/dQ)*(P/Q), e=(.24*1973\^.2)\*(2000/1973)= 1.109

iii. Benefits are unclear for owners of old cars because old does not mean that it is worth less than \$2000. If a car is worth more than 2000 dollars than it would not be beneficial for owners to sell back to the state. But if the old car is worth less than 2000 than it would stand to reason that these cars would come out of the woodwork and off the road. This policy may also supply more money to the used car market, because owners of older cars may be priced out of the new car market and they may take their rebate back into the used car market driving prices up for older cars further limiting the amount of cars available to the state at the price of 2000.

iv. Under current conditions the cost to the state would be 1973\*2000= \$3,946,000

3b.

```{r}
MB <- function(q) {
  p=1000-0.1*q^1.2
  return(p)
}

social_equ <- function(q) {
  z=sup(q)-MB(q)
  return(z)
}

social_benefit <- fzero(social_equ, c(0,6000))$x

social_benefit

MB(716)
sup(716)
```

3b i. The socially optimal number of cars to remove would be 716.

ii. Socially optimal price \$733.24
