---
title: "Rennacker_HW_3"
author: "Travis Rennacker"
format: 
  html: 
    toc: true
    code-fold: true
    code-tools: true 
    code-summary: "Show Code"
    embed-resources: true
execute:
  warning: false 
  message: false
---

Libraries

```{r}
#|output: false 

library(here)
library(janitor)
library(tidyverse)
library(pracma)

```




```{r}
MAC_M <- function(a){
  m=a*4
  return(m)
}

MAC_K <- function(a){
  m=a*4.5
  return(m)
}

MAC_A <- function(a){
  m=a*3
  return(m)
}

MAC_V <- function(a){
  m=a*2
  return(m)
}
```
  Prices are in $/Acre Foot and quantities are in Acre Feet (AF).
â— Suppose groundwater abatement costs for each firm have been estimated for you by a team of
agroecologists and the marginal abatement cost (MAC) curves are as follows:
o Kern: MAC = 4.5A (where A = abatement)
o Mojave: MAC = 4A
o Antelope: MAC = 3A
o Ventura: MAC = 2A


```{r}
s_values <- seq(0, 200, by = 1)  
m_values <- sapply(s_values, MAC_M)
k_values <- sapply(s_values, MAC_K)
a_values <- sapply(s_values, MAC_A)
v_values <- sapply(s_values, MAC_V)

Mdata <- data.frame(a = s_values, y = m_values, type = "Mojave")
Kdata <- data.frame(a = s_values, y = k_values, type = "Kern")
Adata <- data.frame(a = s_values, y = a_values, type = "Antelope")
Vdata <- data.frame(a = s_values, y = v_values, type = "Ventura")

# Combine all the data frames into one
all_data <- rbind(Mdata, Kdata, Adata, Vdata)

# Plot using ggplot2
ggplot(all_data, aes(x = a, y = y, color = type)) +
  geom_line() +
  labs(title = "Water Abatement Costs of Four Counties",
       x = "Abatement",
       y = "Price", color="County") +
  theme_minimal() +
  scale_color_manual(values = c("Mojave" = "red", "Kern" = "gold2", "Antelope" = "brown", "Ventura" = "turquoise"))


```

# Question 1 Response: 

While there is no limit on the ability of each county to abate groundwater usage, there is an increased cost associated respective of location. It would be most cost effective for Ventura County to utilize abatement methods. 

# Question 2 Response: 


# 2 a.

Each District would Abate 50 units.

Amount of Abatement
```{r}



MAC_K(50)
MAC_M(50)
MAC_A(50)
MAC_V(50)

abatement_strict <- data.frame(
  price = c(100, 150, 200, 225),
  county = c("Ventura", "Antelope", "Mojave", "Kern"))

abatement_strict

```

```{r}
#to find totals for area under curve(aka area under lines)
bh <- function(b,h){
  a=(b*h)/2
  return(a)
}

bh(50,100) #ventura
bh(50,150) #antelope
bh(50,200) #mojave
bh(50,225) #kern

abatement_strict |> mutate(total_cost = c(2500, 3750, 5000, 5625))

#2500 + 3750 +5000 + 5625 #Total cost for all firms at 200 AF of abatement

```

Total Cost for all counties: $16875

# 2 b. 

MAC_K = 4.5a  :  A_K = p/4.5
MAC_M = 4a    :  A_M = p/4
MAC_A = 3a    :  A_A = p/3 
MAC_V = 2a    :  A_V = p/2

A_K + A_M + A_A + A_V = Aggregate Abatement

Market is set at 200AF of abatement by government

Aggregate Abatement = 200AF

```{r}
#Function of agg cost
f_AA <- function(x) {
  200 - (x / 4.5 + x / 4 + x / 3 + x / 2)
}

#set interval for root
result <- uniroot(f_AA, c(0, 1000))  

#Show results
result$root

```
# Equilibrium Market Price with cap and trade : $153.19 

```{r}

A_K  <- function(p) {
  a = p/4.5
  return(a)
}


A_M <- function(p) {
  a = p/4
  return(a)
} 

A_A <- function(p) {
 a = p/3 
 return(a)
} 

A_V <- function(p) {
  a = p/2
  return(a)
}

bh <- function(b,h){
  a=(b*h)/2
  return(a)
}


bh(A_K(153.1915) , 153.19)
bh(A_M(153.1915) , 153.19)
bh(A_A(153.1915) , 153.19)
bh(A_V(153.1915) , 153.19)

dfa <- data.frame(abatement = c(A_K(153.1915),A_M(153.1915),A_A(153.1915),A_V(153.1915)), county= c("Kern", "Mojave", "Antelope", "Ventura"), "total abatement cost per county" = c(2607.49, 2933.426, 3911.234, 5866.851))

dfa

(A_K(153.1915) * 153.19)/2 + (A_M(153.1915) * 153.19)/2 + (A_A(153.1915) * 153.19)/2 + (A_V(153.1915) * 153.19)/2

16875 - 15319 


```
The above table displays how much abatement each county would revert to in a cap and trade scenario. 

Total Abatement Cost  :  $ 15319

Total gains under cap and trade :  16875 - 15319 = **$1556**


# 2 c. 

To decrease the amount of water used by 200AF, a tax should be set at $ 153.1915.

```{r}
dfa 

sum(dfa$abatement, na.rm = TRUE)

```
Each district would abate the above amount. Total cost is also listed above. 

Total Abatement Across Districts: 200 AF

Total cost of abatement for each district to comply listed in the above table  

```{r}
#THIS IS WRONG THEY ALL DONT ABATE 100 They each used 100... 
f_AMV <- function(p) {
  100 - ( p / 4 + p / 3 + p / 2)
}

#set interval for root
result <- uniroot(f_AMV, c(0, 1000))  

#Show results
result$root

A_M(92.307)
A_A(92.307)
A_V(92.307)

MAC_K(200)

```

```{r}


# 100 AF for  
# (100-abatement_per_county)*tax
#This would be the taxed water usage for each county (initial water usage - abatement)*tax rate
kern  <- (200 - 34.04256)*153.1915	#Kern			
 mojave <- (100 - 38.29787)*153.1915	#Mojave			
   antelope <- (100 - 51.06383)*153.1915	#Antelope		
    ventura <-(100 - 76.59575)*153.1915	#Ventura

tax_revenue <- data.frame(county= c("Kern", "Mojave", "Antelope", "Ventura"), "tax revenue" = c(kern, mojave, antelope, ventura))

tax_revenue

```



c. Tax on groundwater use. Suppose instead a single tax is implemented on all water use 
such that water use is reduced by 200AF summed across the four districts. What is that 
tax level? How much would each district abate? What is total abatement across 
districts? What is the total cost of abatement for each district to comply? How much tax 
revenue is generated if initial water use is 100 AF for Ventura, Antelope, and Mojave 
and 200 for Kern? 


